\textbf{ Spectrophotometric parallaxes with \project{The Cygnet}:\\ Accurate distances for luminous red-giant stars }

### 2018-08-02 paper/sp.tex
intro paragraphs

\title{\textbf{%
Spectrophotometric parallaxes with \project{The Cygnet}:\\
Accurate estimates of luminosity and distance\\
for luminous red-giant stars
}}

@@

\section{Introduction} \label{sec:intro}

@@

In this contribution we develop data-driven techniques to predict
stellar luminosities and distances with spectroscopy and photometry...

The history and value (and disadvantages)
of data-driven models for these kinds of problems.
How could it be that we are completely independent of stellar models?

The dangers of aggressive machine-learning methods. Why will we just
look at linear models? The disadvantages of this approach.

The dangers of cutting on \gaia\ properties. Why we will not cut on
\gaia\ parallaxes. The disadvantages of this approach.

\section{Why and how should we model red-giant stars?}

HOGG: We need to, for \sdssv.

HOGG: We know we can because of physical arguments.

HOGG: What kinds of inputs do we need to predict parallaxes?

>>>>

\title{\textbf{%
Spectrophotometric parallaxes with \project{The Cygnet}:\\
Accurate distances for luminous red-giant stars
}}

@@

\section*{~}\clearpage
\section{Introduction} \label{sec:intro}

@@

Luminous red-giant stars are valuable for mapping the Milky Way disk for a number
lf reasons, only one of which is their great luminosities (and hence brightnesses
even at large distance).
They are extremely common (frequent) stars, much more common than distance-indicating
stars like Cepheids.
They are produced by stellar populations at all metallicities and almost all ages,
so they can be found in all parts of the Galaxy.
They have temperatures and surface gravities such that it is
straightforward to spectroscopically measure metallicities,
detailed chamical abundances, and stellar parameters, including ages (CITE MARTIG, NESS).
The red-giant branch is photometrically near-orthogonal to reddening vectors by dust,
so they are easy to de-redden or dust-correct.
And finally, in physical models of red giants, their predicted luminosities are simple
functions of composition, surface gravity, temperature, and age.
Thus, if we can measure these properties of red giants, we can (in principle) predict
their luminosities and use them as standard candles.
That is---like type Ia supernovae---red giants are standardizable candles.

In this project, we develop purely data-driven techniques to predict
red-giant parallaxes or distances with spectroscopy and photometry.
Our method capitalizes on the point that red-giant stars are dust-correctable,
standardizable candles, but it makes no use whatsoever of any stellar interior,
stellar exterior, or dust-reddening models.
We only use these physical expectations about red giants to recommend
data features---inputs---for the data-driven model.
The data-driven method uses patterns in the observed data themselves
to discover the relationships between spectral features in the spectra of the stars,
photometry (including colors), and parallax (or distance).

Data-driven models are very often more accurate than physical models of stars
for this kind of work, because they have the flexibility to discover
regularities or offsets in the data that are not captured by the physical
models.
They do better than physical models when there are lots of training data
with properties that are good for performing regressions.

Simple data-driven models (like the ones we will employ here)
suffer from many limitations for this kind of work:
Because they are constrained only by the data, and not physical law,
they can learn relationships that we know
to be physically unlikely, and (by the same token) they can't benefit from our
physical knowledge.
They are required, in some sense, to spend some of the information in the data on
learning physical laws or relationships.
Not all of the information in the data is being used
optimally when we ignore our physical models.
Relatedly, they are not (usually) interpretable or generalizable:
A data-driven model trained with one kind of data can not usually be used in another
context with very different input data.
And the internal properties of the model are not useful, in general, for improving
our understanding or intuition about the astrophysics in play.

There is a huge range of complexity or capacity in data-driven
models, from linear regressions (what we use here) to the most advanced machine-learning
methods (like deep neural networks).
In this work we stay on the low-complexity side of this.
We do not want to face the (literally combinatoric) choices in model structure that
advanced methods bring.
And we want to control or understand the smoothness or flexibility of the model.
Deep networks can model extremely complex function spaces, which is good in some
contexts, but bad when you know that the functional dependencies you hope to
model are smooth:
In that case, much of the information in the data can get spent on learning the
smoothness of the function.

All that said, our goal here is to produce a precise mapping tool for surveys
like \apogee\ and \sdssv.
So we will take the limitations of the data-driven models in exchange for their
performance.
This work builds on the success we have seen in building other kinds of
data-driven tools for  stellar spectroscopy, for example to measure stellar
parameters (CITE NESS), abundances (CITE HO, NESS, CASEY), and masses and ages (CITE NESS).

HOGG: The dangers of cutting on \gaia\ properties. Why we will not cut on
\gaia\ parallaxes. The disadvantages of this approach.
